---
title: "K-means Clustering of Stocks "
author: "Michael Wanek, Woodny Dorceans"
date: '`r Sys.Date()`'
format:
  html:
    code-fold: true
course: STA 6257 - Advance Statistical Modeling
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
---

## Introduction
Through the use of data mining technology, large amounts of complex financial data can be analyzed. K-means clustering (KMC) is an algorithm that can be used for potentially maximizing profit, or reducing risk, when investing in company stock (Soofi et al., 2015). Using KMC, stock data can be grouped together in accordance with predetermined criteria to find similarity, dissimilarity, and structure. KMC can also be used to classify financial features according to maximum and minimum similarity.Clustering algorithms, such as the k-means clustering (KMC) algorithm, have gained attention as valuable tools for aiding investment decision-making. However, there are some shortcomings to the method including the determination of the number of ‘k clusters’, different distance calculation methods, and the problem of local extremum (Fang et al., 2021).

What is k-means clustering(KMC)? It is a version of unsupervised machine learning. Unsupervised learning would utilize algorithms to analyze and cluster datasets that are unlabeled (Malik et al., 2019). In this case, K-means clustering would be considered an unsupervised learning method where similar data points will be assembled into groups of unlabeled data . It groups similar unlabled data by looking at the average distance between the objects in each group which is known as the centroid and the K groups. Then you must measure the distance between each objects and centroid and assigned to the correct groups until all objects have a group (Yuan et al., 2019). In order to weigh variables one process that can be used is the ‘Analytic Hierarchal Process’. The stages of the KMC algorithm were as follows:
1.	Initial stage that partitioned the objects randomly into ‘k’ clusters.
2.	The repetition stage by calculating the center of each cluster using the mean of the data, compute the squared Euclidean distance from each object to each cluster, and compute the squared error function.
3.	The improvement stage where objects were assigned to the cluster with the nearest center.
4.	The stop stage which was a process that continued until no object move clusters or the objective function value doesn’t reduce (Soofi et al., 2015).

Using KMC it can help a stock buyers and sellers understand the stock market pattern. K-means clustering when combined with regression method also helps with predicting stock future stock prices. This allows users to know when the best time to get in the market before a price increase/decrease, which in turn, tell sellers to hold or to sale their stocks (Bini et al., 2016). KMC is overall a great algorithm, but it also does have limitations. Some of those limitations include the determination of the number of ‘k clusters’, different distance calculation methods, and the problem of local extremum (Fang et al., 2021). Other types of limitations include the inability to use all types of data with KMC (Ahmed et al., 2020).


This paper provides the definition of KMC, some of the limitations of the methodology, as well as our analysis on how well KMC does handling stock data. 


## Methods


$$

$$


```{r, warning=FALSE, echo=T, message=FALSE}

```


```{r, warning=FALSE, echo=TRUE}

  
```

### Analysis and Results
```{r, warning=FALSE, echo=TRUE}
library(readr)
SPDataset <- read_csv("SPDataset.csv")
```

### Statistical Modeling

### Conlusion

## References
Ahmed, M., Seraj, R., & Islam, S. M. S. (2020). The k-means algorithm: A comprehensive survey and performance evaluation. Electronics, 9(8), 1295.

Babu, M. S., Geethanjali, N., & Satyanarayana, B. (2012). Clustering approach to stock market prediction. International Journal of Advanced Networking and Applications, 3(4), 1281.

Bini, B. S., & Mathew, T. (2016). Clustering and regression techniques for stock prediction. Procedia Technology, 24, 1248-1255.

Fang, &amp; Chiao, C. (2021). Research on prediction and recommendation of financial stocks based
on K-means clustering algorithm optimization. Journal of Computational Methods in Sciences
and Engineering, 21(5), 1081–1089. https://doi.org/10.3233/JCM-204716

Malik, &amp; Tuckfield, B. (2019). Applied unsupervised learning with R: Uncover hidden
relationships and patterns with K-Means clustering, hierarchical clustering, and PCA (1st
edition). Packt Publishing Ltd.

Ossareh, Pourjafar, M. S., &amp; Kopczewski, T. (2021). Cognitive Biases on the Iran Stock
Exchange: Unsupervised Learning Approach to Examining Feature Bundles in Investors’
Portfolios. Applied Sciences, 11(22), 10916–. https://doi.org/10.3390/app112210916

Soofi , Mohseni , M., &amp; Momeni , M. (2015). Clustering Stock Market Companies via K- Means
Algorithm. Kuwait Chapter of Arabian Journal of Business &amp; Management Review, 4(5), 1–10.
https://doi.org/10.12816/0018959

Yedla, M., Pathakota, S. R., & Srinivasa, T. M. (2010). Enhancing K-means clustering algorithm with improved initial center. International Journal of computer science and information technologies, 1(2), 121-125.

Yuan, C., & Yang, H. (2019). Research on K-value selection method of K-means clustering algorithm. J, 2(2), 226-235.

Zuhroh, Rofik, M., &amp; Echchabi, A. (2021). Banking stock price movement and
macroeconomic indicators: k-means clustering approach. Cogent Business &amp; Management,
8(1), 1–10. https://doi.org/10.1080/23311975.2021.1980247



